Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-03 paper/paper.tex
minor fixups

\subsection{Long-term rewards and action policies}
The behavior of the agent is governed by some kind of \textit{policy},
which maps states of the world to
a set of candidate actions to perform given this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory of action cascades as follows
\begin{eqnarray*}
  \begin{split}
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect rewards }R_0 = r(s_0, a_0)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect rewards }R_1 = r(s_1, a_1)\\
    \ldots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect rewards }R_{t} = r(s_{t}, a_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}
Since an action taken in the present moment $t$
might have repercussions long into the future, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account expected reward in the future.
A common approach to
modeling this accumulation is the time-discounted \textit{cumulative reward}% function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a policy $\pi$.
\begin{mdframed}
  \textbf{Where is ``value'' buffered in the DMN ?
    % brain ?
    }
  The amygdala is known to be involved in significance evaluation while
  the NAc, which is connected to the vmPFC, is involved in reward evaluation.
\end{mdframed}
  The goal of the RL agent is then to adapt this policy so as to  maximize $G^\pi$. In
  the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
Setting $\gamma = 0$ corresponds to a perfectly shortsighted agent who is solely concerned about its immediate rewards.
Such an agent would have no horizon for the future,
which is not compatible with long-term goal planning
as potentially implemented in the DMN.
To allow a learning process to arise,
it is necessary that $0 < \gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
Such an agent puts relatively more emphasis on rewards expected in
a short future and pays less attention to
rewards expected at a later time.
% given that there is more uncertainty in the farsighted future.
The reward discount factor $\gamma$ can be seen as calibrating
the delay discounting behavior of the intelligent agent,
that is, the decision related to reward-delay trade-off schedules.
It is also important to appreciate that a stochastic policy estimation
is more advantageous in many RL settings.
For instance, a deterministic strategy in playing rock-paper-scissors can
be easily exploited and a uniformly random choice is optimal.

@@

Learning
of the the entire model parameters can effectively be achieved
via \textit{backpropagation},
where prediction errors (i.e., \textit{regrets}) percolate from lower to higher processing layers to modulate the choice of future actions
% online stochastic gradient descent is used to minimize the variational objective
% \eqref{eq:oracle}.

% \paragraph{Parameter update.}
\begin{equation}
  \delta \theta_{k+1}^Q \propto -\nabla_{\theta^Q_{k+1}}\mathcal L(\theta^Q_{k+1})
  % = -\mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{}))^2]
  = -\mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret}}
    \underbrace{\nabla_{\theta^Q_{k}}Q(s, a|\theta^Q_{k+1})}_{\text{???}}\right].
  \label{eq:oracle}
\end{equation}

@@

% \begin{figure}[!tb]
%   \centering
%   \includegraphics[width=.8\linewidth]{vae.pdf}
%   \caption{Variational autoencoders...}
%   \label{fig:vae}
% \end{figure}
The main algorithm for minimizing free-energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}.
\citep{dayan1995helmholtz}. As noted already in \citep{dayan1995helmholtz},
a crucial drawback of the wake-sleep algorithm is that it requires a concurrent models (generative and recogntiion), which together do not correspond to optimization of (a bound of) the marginal likelihood (because of the incorrect KL used therein, etc.). Thus the brain could not possible be running such a algorithm, not even in principle! To the rescue, we note that the recent theory of
\textit{variational auto-encoders (VAEs)} \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm, as it overcomes the technical limits of the former, by using a reparametrization trick. For instance, unlike the wake-sleep algorithm for minimizing free-energy, VAEs can be efficiently trained via back-propagation of learning errors.

>>>>

\subsection{Long-term rewards and action policies}
The behavior of the agent is governed by some kind of \textit{policy},
which maps states of the world to
a set of candidate actions to perform given this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory of action cascades as follows
\begin{eqnarray*}
  \begin{split}
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect reward }R_0 = r(s_0, a_0)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect reward }R_1 = r(s_1, a_1)\\
    \ldots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect reward }R_{t} = r(s_{t}, a_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}
Since an action taken in the present moment $t$
might have repercussions long into the future, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account expected reward in the future.
A common approach to
modeling this accumulation is the time-discounted \textit{cumulative reward}% function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a policy $\pi$.
\begin{mdframed}
  \textbf{Where is ``value'' buffered in the DMN ?
    % brain ?
    }
  The amygdala is known to be involved in significance evaluation while
  the NAc, which is connected to the vmPFC, is involved in reward evaluation.
\end{mdframed}
  The goal of the RL agent is then to adapt this policy so as to  maximize $G^\pi$. In
  the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
Setting $\gamma = 0$ corresponds to a perfectly shortsighted agent who is solely concerned about its immediate rewards.
Such an agent would have no horizon for the future,
which is not compatible with long-term goal planning
as potentially implemented in the DMN.
To allow a learning process to arise,
it is necessary that $0 < \gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
Such an agent puts relatively more emphasis on rewards expected in
a short future and pays less attention to
rewards expected at a later time.
% given that there is more uncertainty in the farsighted future.
The reward discount factor $\gamma$ can be seen as calibrating
the delay discounting behavior of the intelligent agent,
that is, the decision related to reward-delay trade-off schedules.
It is also important to appreciate that a stochastic policy estimation
is more advantageous in many RL settings.
For instance, a deterministic strategy in playing rock-paper-scissors can
be easily exploited and a uniformly random choice is optimal.

@@

Learning
of the the entire model parameters can effectively be achieved
via \textit{backpropagation},
where reward prediction errors (RPES) -- also known as \textit{regret}-- percolate from lower to higher processing layers to modulate the choice of future actions
% online stochastic gradient descent is used to minimize the variational objective
% \eqref{eq:oracle}.

% \paragraph{Parameter update.}
\begin{equation}
  \delta \theta_{k+1}^Q \propto -\nabla_{\theta^Q_{k+1}}\mathcal L(\theta^Q_{k+1})
  % = -\mathbb E_{(s, a, r, s') \sim U(\mathcal D)}[(Q(s, a|\theta^Q) - Q(s',a'|\theta^Q_{}))^2]
  = -\mathbb E_{(s, a, r, s') \sim \mathcal U(D)}\left[\underbrace{(Q(s, a|\theta^Q_{k+1}) - y)}_{\text{regret / RPE}}
    \underbrace{\nabla_{\theta^Q_{k}}Q(s, a|\theta^Q_{k+1})}_{\text{averseness}}\right].
  \label{eq:oracle}
\end{equation}

@@

% \begin{figure}[!tb]
%   \centering
%   \includegraphics[width=.8\linewidth]{vae.pdf}
%   \caption{Variational autoencoders...}
%   \label{fig:vae}
% \end{figure}
The main algorithm for minimizing free-energy $F_G^R(\x)$ is the \textit{wake-sleep algorithm}.
\citep{dayan1995helmholtz}. As noted already in \citep{dayan1995helmholtz},
a crucial drawback of the wake-sleep algorithm is that it postulates a pair forward (generation) and backward (recognition) models,
which together do not correspond to optimization of (a bound of) the marginal likelihood, because of the incorrect KL used therein, etc.
Thus the brain could not possibly be running such a algorithm, not even in principle! To the rescue, we note that the recent theory of
\textit{variational auto-encoders (VAEs)} \citep{kingma2013auto} might provide an efficient alternative to the wake-sleep algorithm,
as it overcomes the technical limits of the former, by using a reparametrization trick. For instance, unlike the wake-sleep algorithm
for minimizing free-energy, VAEs can be efficiently trained via back-propagation of learning errors.
